---
title: "hospital-admissions"
author: "George WIlloughby"
date: "10/06/2021"
output: html_document
---

```{r}
library(httr)
library(jsonlite)
library(dplyr)
```

# Getting the hospital data for the NHS Trusts.

```{r}
#' Extracts paginated data by requesting all of the pages
#' and combining the results.
#'
#' @param filters    API filters. See the API documentations for 
#'                   additional information.
#'                   
#' @param structure  Structure parameter. See the API documentations 
#'                   for additional information.
#'                   
#' @return list      Comprehensive list of dictionaries containing all 
#'                   the data for the given ``filter`` and ``structure`.`
get_paginated_data <- function (filters, structure) {
  
    endpoint     <- "https://api.coronavirus.data.gov.uk/v1/data"
    results      <- list()
    current_page <- 1
    
    repeat {

        httr::GET(
            url   = endpoint,
            query = list(
                filters   = paste(filters, collapse = ";"),
                structure = jsonlite::toJSON(structure, auto_unbox = TRUE),
                page      = current_page
            ),
            timeout(10)
        ) -> response
        
        # Handle errors:
        if ( response$status_code >= 400 ) {
            err_msg = httr::http_status(response)
            stop(err_msg)
        } else if ( response$status_code == 204 ) {
            break
        }
        
        # Convert response from binary to JSON:
        json_text <- content(response, "text")
        dt        <- jsonlite::fromJSON(json_text)
        results   <- rbind(results, dt$data)
        
        if ( is.null( dt$pagination$`next` ) ){
            break
        }
        
        current_page <- current_page + 1;

    }
    
    return(results)
    
}


# Create filters:
query_filters <- c(
    "areaType=nhsTrust"
)

# Create the structure as a list or a list of lists:
query_structure <- list(
    date       = "date", 
    name       = "areaName", 
    code       = "areaCode", 
    hosp_cases = "hospitalCases",
    hosp_admissions = "newAdmissions"
)

lower_authority_hosp_cases <- get_paginated_data(query_filters, query_structure)

list(
  "Shape"                = dim(lower_authority_hosp_cases),
  "Data (first 3 items)" = lower_authority_hosp_cases[0:3, 0:-1]
) -> report

print(report)

lower_authority_hosp_cases
```

# See the Liverpool and Sheffield hospitals

```{r}
unique(lower_authority_hosp_cases$name)

```

# Filtering for Liverpool hospitals

```{r}
total_liverpool_hospitals <- filter(lower_authority_hosp_cases, name == "Liverpool Heart and Chest Hospital NHS Foundation Trust" | name == "Liverpool University Hospitals NHS Foundation Trust" | name == "Liverpool Women's NHS Foundation Trust")

# liverpool_hospitals <- rbind(liverpool_heart_chest, liverpool_university_hospitals, liverpool_womens_trust)

total_liverpool_hospitals

```


# Filtering for Sheffield hospitals

```{r}
total_sheffield_hospitals <- filter(lower_authority_hosp_cases, name == "Sheffield Teaching Hospitals NHS Foundation Trust" | name == "Sheffield Health and Social Care NHS Foundation Trust" |name == "Sheffield Children's NHS Foundation Trust")

# sheffield_hospitals <- rbind(sheffield_teaching_hospital, sheffield_health_social_care, sheffield_childrens_hospital)

total_sheffield_hospitals
```

# Removing NAs from the two datasets

Some hospitals had admissions and some didn't which means that there are NA values

```{r}
total_liverpool_hospitals <- total_liverpool_hospitals %>%
  mutate_all(~replace(., is.na(.), 0))

total_sheffield_hospitals <- total_sheffield_hospitals %>%
  mutate_all(~replace(., is.na(.), 0))
```



# Calculating a summary for each of Liverpool and Sheffield hospitals

```{r}
library(dplyr)

sum_total_sheffield_hospitals <- total_sheffield_hospitals %>%
  group_by(date) %>%
  summarise(sum_sheffield_hospitals =sum(hosp_cases))

sum_total_sheffield_hospitals

sum_total_liverpool_hospitals <- total_liverpool_hospitals %>%
  group_by(date) %>%
  summarise(sum_liverpool_hospitals = sum(hosp_cases))

sum_total_liverpool_hospitals
```


# Calculate the hospital admission rates for Liverpool and Sheffield

```{r}
library(dplyr)
library(zoo)

sum_total_liverpool_hospitals_mean <- sum_total_liverpool_hospitals %>%
  dplyr::arrange(desc(date)) %>%
  dplyr::mutate(hosp_cases_seven_day_average_liverpool = zoo::rollmean(sum_liverpool_hospitals, k = 7, align="left", fill = NA)) %>%
  dplyr::ungroup()

sum_total_liverpool_hospitals_mean


sum_total_sheffield_hospitals_mean <- sum_total_sheffield_hospitals %>%
  dplyr::arrange(desc(date)) %>%
  dplyr::mutate(hosp_cases_seven_day_average_sheffield = zoo::rollmean(sum_sheffield_hospitals, k = 7, align="left", fill = NA)) %>%
  ungroup()

sum_total_sheffield_hospitals_mean
```


# Joining Liverpool and Sheffield hospitals

```{r}
combined_liverpool_sheffield_hospitals <- merge(sum_total_liverpool_hospitals_mean, sum_total_sheffield_hospitals_mean, by ="date")

combined_liverpool_sheffield_hospitals
```

# Export the data

```{r}
write.csv(combined_liverpool_sheffield_hospitals, "hospital-admissions-liv-shef.csv") 
```






